# 1. Load necessary libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(car)
library(e1071)
```

# 2. Load the data
### Load the data from a TSV (Tab-Separated Values) file into a data frame named data

```{r}
data <- read_tsv("Clean_data/PlateletHW.tsv")
```

# 3. Data Cleaning
### Clean the data by taking the absolute value of ADP

```{r}
data <- data %>% mutate(ADP = abs(ADP))
```

# 4. Checking for Skewness and Outliers
### Calculate Skewness

```{r}
adp_skewness <- skewness(data$ADP)
```

### Calculate Quartiles and IQR

```{r}
Q1 <- quantile(data$ADP, 0.25)
Q3 <- quantile(data$ADP, 0.75)
IQR <- Q3 - Q1
```

### Determine Outlier Boundaries

```{r}
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
```

### Filter Outliers

```{r}
outliers <- data %>% filter(ADP < lower_bound | ADP > upper_bound)
cat("Skewness of ADP:", adp_skewness, "\n")
cat("Number of outliers:", nrow(outliers), "\n")
```

### Write a new clean data

```{r}
write_tsv(data, "Clean_data/PlateletHW_clean.tsv")
```

# 5. Normalizing ADP by Taking the Logarithm

```{r}
data <- data %>% mutate(ADP_log = log(ADP))
```

