# Creating Adjusted Models with

```{r}
model1_adj <- lm(ADP_log ~ rs4244285 + AGE + SEX, data = data)
model2_adj <- lm(ADP_log ~ rs4986893 + AGE + SEX, data = data)
model3_adj <- lm(ADP_log ~ rs662 + AGE + SEX, data = data)

logit_model1_adj <- glm(Resistance ~ rs4244285 + AGE + SEX, data = data, family = binomial)
logit_model2_adj <- glm(Resistance ~ rs4986893 + AGE + SEX, data = data, family = binomial)
logit_model3_adj <- glm(Resistance ~ rs662 + AGE + SEX, data = data, family = binomial)
```

# Summary of Adjusted Linear Regression Model 1: ADP_log ~ rs4244285 + AGE + SEX

```{r}
summary(model1_adj)
```

# Summary of Adjusted Linear Regression Model 2: ADP_log ~ rs4986893 + AGE + SEX

```{r}
summary(model2_adj)
```

# Summary of Adjusted Linear Regression Model 3: ADP_log ~ rs662 + AGE + SEX

```{r}
summary(model3_adj)
```

# Summary of Adjusted Logistic Regression Model 1: Resistance ~ rs4244285 + AGE + SEX

```{r}
summary(logit_model1_adj)
```

# Summary of Adjusted Logistic Regression Model 2: Resistance ~ rs4986893 + AGE + SEX

```{r}
summary(logit_model2_adj)
```

# Summary of Adjusted Logistic Regression Model 3: Resistance ~ rs662 + AGE + SEX

```{r}
summary(logit_model3_adj)
```
